---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [Nearly 250 women have been
fatally shot by police since 2015](https://www.washingtonpost.com/graphics/2020/investigations/police-shootings-women/)

## Load libraries
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
```
Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
fatal_police_shootings <-read_csv("fatal-police-shootings-data.csv")
fatal_police_shootings_clean <- read_csv("fatal-police-shootings-data-clean.csv")



```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: ["Of men killed, 57 percent were armed with a gun and of women, 44 percent. In other cases, women were armed with knives, cars, toy weapons, hammers and hatchets."
* **Analysis summary**: We used open refine to clean up the gun column and get all of the incidents with guns. We then created separate data frames for each gender with incidents involving a gun. Then we calculated the percentages. Our results were not identical to the Post's and this may have had to do with them calculating it at a different time, but our results are close.
]

```{r}
all_police_shootings_by_gender <- fatal_police_shootings_clean%>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
all_shootings_men <- all_police_shootings_by_gender %>% 
  filter(gender == 'M')
all_shootings_women <- all_police_shootings_by_gender %>% 
  filter(gender == 'F')
gun_police_shootings_by_gender <- fatal_police_shootings_clean %>%
  filter(armed_clean == "gun") %>%
  group_by(gender) %>%
  summarise(count=n()) %>%
  arrange(desc(count))
  
all_shootings_men_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'M')
all_shootings_women_with_gun <- gun_police_shootings_by_gender %>% 
  filter(gender == 'F')
men_with_gun_percentage <- ((4012/6814)*100)
women_with_gun_percentage <- ((145/320)*100)
men_with_gun_percentage
women_with_gun_percentage


```

### Sentence 2

* **Sentence text**: [“Since 2015, police have killed 26 unarmed women, including Taylor. Of those, 14 were White, seven were Black, four were Hispanic and one’s race was unknown. While about twice as many White women were shot and killed as Black women, White women account for five times the Black female population.” ]
* **Analysis summary**: [* **Analysis summary**: We've managed to replicate the Post's findings almost. However, we have different totals for white, Hispanic, and other races than they did. We get closer if we filter by age, but that inst still exact. We don't know what age they were defining as making someone a woman.

]

```{r}
shootings_women_race<-fatal_police_shootings%>%
  filter(armed == "unarmed", gender=="F", age >17)%>%
  group_by(armed, gender, race)%>%
  summarise(
    count=n()
    )

```

### Sentence 3

* **Sentence text**: ["Since 2015, Black women have accounted for less than 1 percent of the overall fatal shootings in cases where race was known. "]
* **Analysis summary**: [We removed all of the N/A results from race. Then we filtered by race and gender and sumarrized to get the total of black women shot by police which is 54. Then we calculated the total shot which is 5941. We got 0.9% which replicates the Post's <1%.It is possible our result is higher if it has been changed since the initial result, but it's still in the given range.


```{r}
shootings_race_known<-fatal_police_shootings%>%
 drop_na(race)
  shootings_black_women<-shootings_race_known%>%
  filter(gender == 'F', race == 'B' )%>%
  summarise(
    count=n()
    )
  
shootings_race_known%>%
   summarise(
    count=n()
    )
percent_black_women<- 
  ((54/5941)*100)
```

### Sentence 4

* **Sentence text**: [Black men, 12 percent of the male population, make up 27 percent of the men shot and 36 percent of the unarmed deaths.]
* **Analysis summary**: []

```{r}
shootings_race_known<-fatal_police_Shootings %>% 
drop_na(race)

shootings_women<-shootings_race_known %>% 
  filter(gender=="F") %>% 
  summarise(
    count=n()
  )

percent_black_women<-
  
  ((54/280)*100)

shootings_unarmed_women<-fatal_police_Shootings %>% 
  filter(armed=="unarmed", gender=="F") %>% 
  summarise(
    count=n()
  )
percent_black_women_unarmed <- 
  ((7/27)*100)
```

### Sentence 5

* **Sentence text**: [Paste in sentence to engineer here]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
shootings_race_known<-fatal_police_shootings %>% 
  drop_na(race)

shootings_men<-shootings_race_known %>% 
  filter(gender=="M") %>% 
  sumarise(
    count=n()
  )

shootings_black_men<-shootings_race_known %>% 
  filter(gender=='M', race=='B') %>% 
  summarise(
    count=n()
  )
percent_black_men<-
  ((1538/5660)*100)

Shootings_unarmed_men<-fatal_police_shootings %>% 
  filter(armed=="unarmed", gender=="M") %>% 
  group_by(armed,gender,race) %>% 
  summarise(
    count=n()
  )
percent_unarmed_black_men<-
  ((133/389)*100)
```

-30-